---
title: "Lab 5"
author: "John Chandara"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```


```{r}
accident <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```


1. Are there some days of the week where more accidents happen than on others (use variable `DAY_WEEK`)?
```{r}
  accident %>%
    ggplot(aes(x=wday(DAY_WEEK, label=TRUE))) + geom_bar() + xlab('Day of the Week') + ylab('Accidents')
```

2. What time of the day do accidents happen (use variable `HOUR`)?
```{r}
  accident %>% filter(HOUR < 24) %>%
    ggplot(aes(x=HOUR)) + geom_bar() +
    scale_x_discrete(labels=c(
      "0" = "12:00 AM",
      "5" = "05:00 AM",
      "10" = "10:00 AM",
      "15" = "03:00 PM",
      "20" = "08:00 PM"
      )) +
    xlab('Hour of the Day') + ylab('Accidents')
```

3. What is the number of accidents with at least one drunk driver (use variable `DRUNK_DR`)?

```{r}
  result <- accident %>% filter(DRUNK_DR >= 1) %>%
    summarize(n = n())

  result$n
```

There exist 8769 accidents in this data set with at least one drunk driver.

8. Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.
