---
title: "Lab 5"
author: "Michael Egle"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```


```{r}
accident <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```


1. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?


It appears that weekends (Friday, Saturday, and Sunday) tend to have the most accidents

2. What time of the day do accidents happen (use variable HOUR)?


There's a small spike at about 6am when people could be making their daily commute to work. Accidents also gradually increase to a peak at 6pm.

3. What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?


8769 of the accidents had at least one drunk driver involved.

4. In the person table, identify drivers (PER_TYP == 1, see FARS manual) and subset on them.

```{r}
drivers <- person %>%
  filter(PER_TYP == 1)
```

5. Join accident and person table (work out which variable(s) to use)

```{r}
accident <- accident %>%
  inner_join(person)
```

6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

```{r}
accident %>%
  group_by(DAY_WEEK, HOUR, SEX) %>%
  filter(HOUR != 99, SEX == 1 | SEX == 2) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = HOUR, y = count, fill = SEX)) +
  facet_wrap(. ~ DAY_WEEK) +
  geom_bar(stat = "identity")
```


7. Now plot a choropleth map of accidents on a county level

Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs

```{r}
fars <- readxl::read_xlsx('GLC.xlsx')
```
```{r}
fars <- fars %>%
  mutate_at(.vars = vars(`State Code`, `County Code`),
            .funs = funs(as.integer))

accident <- accident %>%
  inner_join(fars, by = c('STATE' = 'State Code', 'COUNTY' = 'County Code'))
```


8. Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.

9. Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?
